---
title: "ggplot()"
subtitle: "FSH 507 Fall 2019"
author: "Amanda Warlick"
date: " `r Sys.Date()`"
fontsize: 10pt
output:
  # powerpoint_presentation: default
  ioslides_presentation: default
  html_document: default
  # slidy_presentation: default
---

```{r setup, include=FALSE, echo = F}

knitr::opts_chunk$set(warning = F, echo = T, tidy = T, fig.height = 3, fig.width = 4, fig.align = 'center', 
                      tidy.opts = list(width.cutoff=10), width = 10)

library(tidyverse)
library(gapminder)
library(dplyr)
library(ggplot2)
library(ggmap)
library(maps)
library(mapdata)
library(cowplot)
library(gganimate)


#the dreaded setwd()
setwd("~/Documents/SAFS/FSH507R_ggplot/presentations/For Lecture")
uspop_data <- read.csv('USpop.csv', header = T, stringsAsFactors = F) %>%
  transform(region = tolower(region))
data <- read.csv('PiGu_exDat.csv', header = T, stringsAsFactors = F)

data_means <- data %>%
  group_by(year, mo, site) %>%
  summarize(temp = mean(temp, na.rm = T), count = mean(PG_count), tide = mean(v))

mycolors <- c("violetred4", "dodgerblue3", 'deepskyblue1', "#a3d39c", "#f6b61c", "chocolate2", "red3")

my_theme <- function(...) {
  theme(legend.title = element_blank(), #play around with turning these elements on and off one at a time
        plot.background = element_rect(), 
        panel.background = element_rect(fill = 'white'), #color background of plot
        panel.border = element_rect(fill = NA), #border of plot
        panel.grid = element_blank(),
        legend.key = element_blank(),
  ...)
}

```

## Benefits of ggplot()

- Can do everything (and more?!) that base R plotting can do
- Tidyverse - streamlined, quick modifications, data-subsetting
- Many extensions, color themes, and (easy) customization options

## The Basics

- Data
- Plot type: geom_*
- Aesthetics: aes(x, y, grouping)
- Scales: axes, colors
- Appearance using theme()
- Customizations: axes, fonts, legends

## The Basics

- Data
    - Dataframes by columns and grouping variables  
    - Seabird counts by site, year, month, yday and temperature and tide   
    - Covariates: temperature, tide height, minutes since high tide  

```{r, echo = TRUE, size = 1}

head(data)

```

## Scatterplot

- Plot type
    - Controlled by 'geom_*()' argument

```{r }

ggplot(data, aes(yday, temp)) + 
  geom_point()

```

## Customize
- Axis labels  
- aes(): change color OUTSIDE aes() to manually designate

```{r }

ggplot(data, aes(yday, temp)) + 
  geom_point(color = 'blue') +
  xlab('Year Day') + ylab('Temperature') +
  theme_classic() #other simple versions include theme_bw()

```

## Customize
- Axis labels  
- aes(): change color INSIDE aes() to group by variable

```{r }

ggplot(data, aes(yday, temp)) + 
  geom_point(aes(color = factor(mo))) +
  xlab('Year Day') + ylab('Temperature') +
  theme_classic() #other simple versions include theme_bw()

```

## Customize 
- Change colors, sizes, and legend labels
- Overlapping points: alpha for transparency or position = 'jitter'

```{r }

ggplot(data, aes(yday, temp)) + 
      #try grouping color by year or site
  geom_point(aes(color = factor(mo))) + 
  #try adding size = year inside aes()
  # geom_point(aes(color = factor(mo), size = year, alpha = 0.2), show.legend = F) +
  xlab('Year Day') + ylab('Temperature') +
  theme(legend.title = element_blank(), #play around with turning these elements on and off one at a time
        plot.background = element_rect(), 
        panel.background = element_rect(fill = 'white'), #color background of plot
        panel.border = element_rect(fill = NA), #border of plot
        panel.grid = element_blank(),
        legend.key = element_blank(),
        legend.position = 'top') +
  scale_color_manual(labels = c('May', 'June', 'July', 'August'), values = mycolors) 
  
```

## Customize - class exercise
- Play with elements of theme():
    - [Elements list](https://ggplot2.tidyverse.org/reference/theme.html)  
    - [Designed themes](https://ggplot2.tidyverse.org/reference/ggtheme.html)
    - Then check out my_theme() in setup chunk

```{r }

ggplot(data, aes(yday, temp)) + 
  geom_point(aes(color = factor(mo))) + 
  xlab('Year Day') + ylab('Temperature') +
  theme(legend.title = element_blank(), #play around with turning these elements on and off one at a time
        plot.background = element_rect(), 
        panel.background = element_rect(fill = 'white'), #color background of plot
        panel.border = element_rect(fill = NA), #border of plot
        panel.grid = element_blank(),
        legend.key = element_blank(),
        legend.position = 'top') +
  scale_color_manual(labels = c('May', 'June', 'July', 'August'), values = mycolors)
  

```

## Other plot types
- Boxplot

```{r }

ggplot(data, aes(factor(mo), temp)) + 
  geom_boxplot(aes(color = factor(mo))) + #try geom_violin()!
  xlab('Month') + ylab('Temperature') +
  my_theme(legend.position = 'top') +
  scale_color_manual(labels = c('May', 'June', 'July', 'August'), values = mycolors)

```

## Faceting
- Common options include moving the strip title, having differing y-axis scales, ncol/nrow
- Facet_wrap versus facet_grid

```{r }

ggplot(data, aes(factor(mo), temp)) + 
  geom_boxplot(aes(color = factor(mo))) + 
  xlab('Month') + ylab('Temperature') +
  my_theme(legend.position = 'top',
           strip.background = element_blank()) +
  facet_wrap(~year) +
  scale_color_manual(labels = c('May', 'June', 'July', 'August'), values = mycolors)

```


## Multiple geom layers

```{r }

ggplot(data_means %>% filter(site == 'Cliffside'), 
       aes(factor(year), temp, col = factor(mo), group = factor(mo))) + 
  geom_point() + 
  geom_line() +
  xlab('') + ylab('Temperature') +
  my_theme(legend.position = 'top',
           strip.background = element_blank()) +
  scale_color_manual(labels = c('May', 'June', 'July', 'August'), values = mycolors)

```

## Stat layers

- geom_smooth()

```{r }

ggplot(data_means %>% filter(site == 'Cliffside'), #look at different sites, or facet by site below
       aes(factor(year), temp, col = factor(mo), group = factor(mo))) + 
  geom_point(size = 0.8) + 
  geom_line(linetype = 'dashed') + #se = F, method = 'lm'
  geom_smooth(aes(fill = factor(mo)), alpha = 0.2, show.legend = F) +
  #or, instead, use geom_ribbon() if have own 95% CI
  #geom_ribbon(aes(mean = mean, ymin = lower, ymax = upper)) +
  xlab('') + ylab('Temperature') +
  # facet_wrap(~site) +
  my_theme(legend.position = 'top') +
  scale_color_manual(labels = c('May', 'June', 'July', 'August'), values = mycolors) +
  scale_fill_manual(values = mycolors)

```

## Stat layers

```{r }

ggplot(data_means, 
       aes(tide, count, group = year)) + 
  geom_point(size = 0.8) +
  # geom_line(linetype = 'dashed') + #se = F, method = 'lm'
  geom_smooth(method = 'lm') +
  xlab('Tide Height') + ylab('Seabird count') +
  facet_wrap(~year, scales = 'free_y')+
  my_theme(legend.position = 'top',
           strip.background = element_blank()) +
  scale_fill_manual(values = mycolors)

```

## Complications

- Sometimes things get ugly with multiple data sources and geoms
    - Using inherit.aes = F, show.legend
    - Being explicit with data = X, aes(), color, etc.


## gganimate for gifs

```{r}

ggplot(gapminder, aes(gdpPercap, lifeExp, size = pop, colour = country)) +
  geom_point(alpha = 0.7, show.legend = FALSE) +
  scale_colour_manual(values = country_colors) +
  scale_size(range = c(2, 12)) +
  scale_x_log10() +
  facet_wrap(~continent) +
  # gganimate parts
  labs(title = 'Year: {frame_time}', x = 'GDP per capita', y = 'life expectancy') +
  transition_time(year) +
  ease_aes('linear')

```

## Map-making
- Birds example

```{r}

pnw_state_outline <- map_data("state", region = c("oregon", "washington")) 

ggplot(pnw_state_outline) +
geom_polygon(aes(x = long, y = lat, group = group), fill = "grey93", color = "grey50", size = 0.2) + 
  coord_fixed(1, ylim = c(47, 49.15), xlim = c(-125, -121)) + #closer in
# coord_fixed(1, ylim = c(45, 49.15), xlim = c(-125, -116)) + #adjust these to zoom in/out; whole state
  geom_point(data = data, aes(long, lat)) +
  xlab('') + ylab('') +
  theme_classic() +
   theme(
    axis.line = element_blank(), 
    # axis.title = element_blank(),  #put these two lines in if don't want axis labels/ticks
    # axis.ticks = element_blank()
    strip.text = element_text(size = 8),
    strip.background = element_blank(),
    legend.title = element_blank(),
    legend.text = element_text(size = 16),
    panel.border = element_rect(colour = "black", fill = NA, size = 1))

```

## Map-making
- US population

```{r}

us_pop <- map_data('state') %>%
  merge(uspop_data, by = 'region')

ggplot(us_pop, aes(long, lat, group = group)) +
  geom_polygon(aes(fill = population)) +
  coord_fixed(1.3) + scale_fill_gradient(trans = 'log10') 

```


## Exercise: explore density plots

- [This example](https://wilkelab.org/practicalgg/articles/health_status.html)


## ggplotly()?
- Something here?

## Power of `cowplot`

```{r}

a <- ggplot(us_pop, aes(long, lat, group = group)) +
  geom_polygon(aes(fill = population)) +
  coord_fixed(1.3) + scale_fill_gradient(trans = 'log10') 

b <- ggplot(data_means %>% filter(site == 'Cliffside'), 
       aes(factor(year), temp, col = factor(mo), group = factor(mo))) + 
  geom_point() + 
  geom_line() +
  xlab('') + ylab('Temperature') +
  my_theme(legend.position = 'top',
           strip.background = element_blank()) +
  scale_color_manual(labels = c('May', 'June', 'July', 'August'), values = mycolors)

plot_grid(a, b, ncol = 1)

```

  






## Resources

The [ggplot2 cheatsheet](https://github.com/rstudio/cheatsheets/blob/master/data-visualization-2.1.pdf)

Examples of manipulating [everything](https://cedricscherer.netlify.com/2019/08/05/a-ggplot2-tutorial-for-beautiful-plotting-in-r/)

A [master list](http://r-statistics.co/Top50-Ggplot2-Visualizations-MasterList-R-Code.html) of (sometimes ugly, but mind-opening) ggplots

Helpful outline of functionality per [specific geoms] (https://ggplot2.tidyverse.org/reference/)

Examples from [R for Data Science](https://r4ds.had.co.nz/data-visualisation.html)

[Nice color palettes](https://ggplot2.tidyverse.org/reference/scale_brewer.html)





