---
title: "Topic"
subtitle: "FSH 507 Fall 2019"
author: "Your Name"
date: " `r Sys.Date()`"
output:
  ioslides_presentation: default
  powerpoint_presentation: default
  html_document: default
  slidy_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(gapminder)
```

# R Markdown

## Getting Started

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown for presentations see <https://bookdown.org/yihui/rmarkdown/presentations.html>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## Slide with R Output

```{r cars, echo = TRUE}
summary(cars)
```

## Slide with Plot

```{r pressure}
pressure %>% 
  ggplot(aes(temperature, pressure)) + 
  geom_point()

```


# Materials


## Suggested Topics

* understanding the philosophy of the grammar of graphics
  - mapping data to aesthetics, displaying aesthetics with geometries

* Messing with attributes of one geometry
  - what goes into `aes` and what goes outside
  - adding aes() inside `ggplot` call vs inside `geom_XX` call

* Different geometries for different kinds of data
  - discrete vs. continuous etc. 
  
* Themes!
  - basic structure of `theme` argument (e.g. element_text, element_rect, etc.)
  - using themes customize appearance of plots
  - change fonts, colors, sizes, etc.

## Suggested Topics

* Faceting 
  * facet_wrap vs facet_grid
  * customizing facets with grid

* Layering multiple geometries /plots
  * discrete + continuous layers
  * controlling aesthetics across multiple layers
  * Multiple plots with patchwork/cowplot
  
* Exporting ggplot objects
  * saving as .rds
  * explorting to pdf/png/tiff etc. 
  
* ggplot2 extensions
  * gganimate
  * ggplotly
  * the list goes one


## Exercise Ideas

* Recreate the most complicated plot you've ever made in ggplot2

* Use ggplot2 to make a first draft of a figure you're working on for a publication/dissertation

* design and apply a custom theme

## Helpful things

I've put a very old seminar I did on ggplot in the presentations folder

the [ggplot2 cheatsheet](https://github.com/rstudio/cheatsheets/blob/master/data-visualization-2.1.pdf) is great!

https://github.com/eco-data-science/VisualizingData/blob/master/ggplot2_intro.Rmd

https://r4ds.had.co.nz/data-visualisation.html


https://towardsdatascience.com/murdering-a-legendary-data-story-what-can-we-learn-from-a-grammar-of-graphics-ad6ca42f5e30

https://cedricscherer.netlify.com/2019/08/05/a-ggplot2-tutorial-for-beautiful-plotting-in-r

https://socviz.co/index.html#preface

